---
title: "Low-Dimensinal Covariance Regression Applied to HCP Data"
output: html_document
author: Christof Seiler
date: "`r Sys.Date()`"
params:
  num_regions: "15"
  num_subjects: "Inf"
  tp_per_subject: "Auto"
  num_samples: 100
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

Apply **low-dimensional** covariance regression model to preprocessed timeseries data from the [Human Connectome Project](https://db.humanconnectome.org/data/projects/HCP_1200).

```{r}
library(CovRegFC)
library(cowplot)
library(ggplot2)
library(reshape2)
library(magrittr)
library(stringr)
library(devtools)
library(rstan)
library(abind)
library(png)
library(cowplot)
library(BatchJobs)
library(coda)
```

## Load HCP Data

Analysis based on data from all 820 subjects in the S900 data release having four complete rfMRI runs (with 100% of collected timepoints), yielding the following outputs at 6 ICA dimensionalities: Group-average parcellations yielded by group-ICA, subject-specific parcellations, subject-specific node timeseries, and a set of subject-specific parcellated connectomes. 

Load subject information and behavioral measurements.

```{r}
subject_info = read.csv("HCP_PTN820/sample_info.csv")
subject_info = subject_info[,c(
  "Subject",
  "Age",
  "Gender",
  "Acquisition",
  "PSQI_AmtSleep", # amount of sleep in hours
  "PSQI_Score" # Pittsburgh Sleep Quality Index (PSQI) Completed
  )] 
```

Classificatoin of sleep duration from Hirshkowitz et al. (2015)^[Hirshkowitz et al. (2015), National Sleep Foundation's sleep recommendations: Methodology and results summary, Sleep Health.]:

* short sleepers: average equal or less than 6 hours each night
* conventional sleepers: average between 7 and 9 hours each night

```{r}
sleep_duration = rep("undefined",nrow(subject_info))
sleep_duration[subject_info$PSQI_AmtSleep <= 6] = "short"
sleep_duration[(subject_info$PSQI_AmtSleep >= 7) & (subject_info$PSQI_AmtSleep <= 9)] = "conventional"
subject_info$sleep_duration = factor(sleep_duration)
table(subject_info$sleep_duration)
str(subject_info)
```

Load timeseries.

```{r}
num_regions = as.integer(params$num_regions)
num_regions
channel_names = paste0("R",1:num_regions)
path = paste0("HCP_PTN820/node_timeseries/3T_HCP820_MSMAll_d",num_regions,"_ts2")
file_names = list.files(path = path,pattern = ".txt")
file_subject_ids = strsplit(file_names,split = ".txt") %>% unlist
ts = lapply(file_subject_ids,function(Subject) {
  print(paste("reading subject:",Subject))
  full_path = paste0(path,"/",Subject,".txt")
  timeseries = read.csv(full_path,header = FALSE,sep = " ")
  timeseries$Subject = Subject
  timeseries
}) %>% do.call(rbind,.) %>% data.frame
names(ts)[1:num_regions] = channel_names
```

Merge timeseries and subject info into one data frame.

```{r}
ts_subject_info = merge(ts,subject_info,by = "Subject")
```

Plot timeseries for one subject.

```{r}
subject_names = names(table(ts_subject_info$Subject))
timeseries = subset(ts_subject_info,Subject == subject_names[1])
timeseries$timepoint = rep(1:(nrow(timeseries)/4),4)
timeseries$run = lapply(paste0("run_",1:4),function(i) rep(i,nrow(timeseries)/4)) %>% unlist
timeseries_long = melt(timeseries,id.vars = c("timepoint","run",names(subject_info)))
ggplot(timeseries_long,aes(x = timepoint,y = value,color = variable)) + 
  geom_line() + facet_wrap(~run)
timeseries_long_subset = subset(timeseries_long,
                                timepoint < 51 & (variable == "R1" | variable == "R2" | variable == "R3"))
ggplot(timeseries_long_subset,aes(x = timepoint,y = value,color = variable)) + 
  geom_line() + facet_wrap(~run)
```

What is the effective samples size computed using autocorrelation function.

```{r}
run = cut_number(1:nrow(timeseries), 4, labels=FALSE)
ess = lapply(subject_names,function(subject_name) {
  sapply(1:4,function(i) {
    timeseries = subset(ts_subject_info,Subject == subject_name)
    Y = timeseries[run==i,channel_names]
    effectiveSize(Y) %>% min
  })  %>% sum
}) %>% unlist
df = data.frame(subject_id=1:length(ess),ess)
ggplot(df,aes(ess)) + geom_histogram()
tp_per_subject = round(min(df$ess))
tp_per_subject
```

Subsample time points within each subject to account for dependencies.

```{r}
#tp_per_subject = as.integer(params$tp_per_subject)
#tp_per_subject
ts_subject_info_subset = subset(ts_subject_info,
                                sleep_duration == "short" | sleep_duration == "conventional") %>% droplevels
subject_names = names(table(ts_subject_info_subset$Subject))
subsample_ids = lapply(subject_names,function(subject_name) {
  ids = which(ts_subject_info_subset$Subject == subject_name)
  seq(from = min(ids),to = max(ids),length.out = tp_per_subject) %>% floor
}) %>% unlist
ts_subject_info_subset = ts_subject_info_subset[subsample_ids,] %>% droplevels
str(ts_subject_info_subset)
table(ts_subject_info_subset$Subject)
ct = table(ts_subject_info_subset[,c("Acquisition","sleep_duration")])/tp_per_subject
addmargins(ct)
```

Subsample subject for power curve.

```{r}
subsample_subjects = function(ts_subject_info_subset,
                              num_subjects,
                              seed = 934321) {
  set.seed(seed)
  total_samples = table(ts_subject_info_subset$sleep_duration)/table(ts_subject_info_subset$Subject)[1]
  num_subject_short = ceiling(num_subjects * total_samples["short"] / sum(total_samples))
  num_subject_conventional = num_subjects-num_subject_short
  short_names = subset(ts_subject_info_subset,sleep_duration == "short")$Subject %>%
    table %>% names %>% sample(.,num_subject_short)
  conventional_names = subset(ts_subject_info_subset,sleep_duration == "conventional")$Subject %>%
    table %>% names %>% sample(.,num_subject_conventional)
  subject_names = c(short_names,conventional_names)
  ts_subject_info_subset[ts_subject_info_subset$Subject %in% subject_names,]
}
```

## Low-Dimensional Covariance Modeling

We can account for correlation between the error terms of the multivariate response variable using covariance regression model. This model is very similar to "seemingly unrelated regressions" (see Zellner^[Zellner (1962), An efficient method of estimating seemingly unrelated regression equations and tests for aggregation bias, JASA.] and Greene^[Greene (2011), Econometric Analysis.]) with an additional random effects term. Repeat many times for power curve.

```{r}
condition = "sleep_duration"
ts_subject_info_subset = within(ts_subject_info_subset,
                                sleep_duration <- relevel(sleep_duration, ref = "conventional"))
num_subjects = as.integer(params$num_subjects)
num_samples = as.integer(params$num_samples)
log_name = paste0("HCP_Restricted_",num_regions,"_",num_subjects,"_",tp_per_subject)
reg = makeRegistry(id = log_name, file.dir = log_name)
batchExport(reg, li = list(channel_names = channel_names,
                           condition = condition,
                           overwrite = TRUE))
fun = function(data) {
  library(CovRegFC)
  CovRegFC::covreg_low(data = data,
                       response = channel_names,
                       condition = condition,
                       participant = "Subject")
}
if(is.na(num_subjects)) {
  batchMap(reg, fun, ts_subject_info_subset)
} else {
  data = lapply(1:num_samples,function(seed) {
    subsample_subjects(ts_subject_info_subset,
                       num_subjects = num_subjects,
                       seed = seed)
  })
  batchMap(reg, fun, data)
}
submitJobs(reg)
waitForJobs(reg,sleep = 360)
fit_list = reduceResultsList(reg)
save(fit_list,file = paste0("fit_list_HCP_reduced_",num_regions,"_",num_subjects,"_",tp_per_subject,".Rdata"))
fit = fit_list[[1]]
```

Traceplots to assess convergence of HMC sampler.

```{r fig.height=5, fig.width=10}
CovRegFC::plot_diagnostics(fit,par_name = "sigma",num_par = 8)
CovRegFC::plot_diagnostics(fit,par_name = "B",num_par = 8)
```

## Visualize Results

Plot regions.

```{r}
path = paste0("HCP_PTN820/groupICA/groupICA_3T_HCP820_MSMAll_d",num_regions,".ica/melodic_IC_sum.sum")
pngs = list.files(path = path,pattern = ".png")
full_paths = paste(path,pngs,sep = "/")
ps = lapply(1:length(full_paths),function(i) 
  CovRegFC::plot_brain(full_paths[i],title = paste0("R",i))
)
p_brains = do.call(plot_grid,c(ps,nrow = 2))
p_brains
ggsave(p_brains,filename = paste0("brains_",num_regions,".pdf"),width = 9,height = 3)
```

Analyze posterior covariance **explained** by simulation conditions.

```{r fig.height=5, fig.width=10}
condition_names = ts_subject_info_subset[,condition] %>% contrasts %>% rownames %>% rev
title = paste(condition_names,collapse = " vs. ")
p_brain_coeff = CovRegFC::plot_coeff(fit,
                                     response = channel_names,
                                     alpha = 0.05/length(channel_names),
                                     title = title,
                                     brain_slices = full_paths)
p_brain_coeff
ggsave(p_brain_coeff,filename = paste0("brains_coeff_",num_regions,"_",num_subjects,"_",tp_per_subject,".pdf"),
       height = 5,width = 10)
```

## Session Info

```{r}
session_info()
```
